\section{stuff that hyperref et all patches}
\subsection{nameref}

\label to store \@currentlabelname and \@currentHref

\slabel (?) similar

\@writetocindents@ (?)

\@sect, \@ssect: To get the content (\NR@gettitle{#7,}\NR@gettitle{#5}) and to add 
 \Sectionformat{#8}
 
\@part, \@spart, \@chapter,  
\M@sect (memoir?) to get the content 

\LT@c@ption: to get the content

\descriptionlabel: to get the content

\ttl@sect@i: (titlesec): to get the content

(beamer): patches section, subsection, subsubsection, theorem begin.

@opargbegintheorem (?): to get content

(amsthm): \@begintheorem to get the name

\ref, \pageref (if hyperref is loaded) to accomodate to the larger number of arguments of label.

(ifthen):\ifthenelse (for some internal problem ...)

\@@vpageref: adapt to changed \newlabel arguments.

(listings): \lst@MakeCaption: to get the caption contents.

 
\subsection{hyperref driver}
in the driver files

\@ssect: to set an anchor

\@schapter:  to set an anchor

\@chapter

\@part

\@spart

\subsection{hyperref.sty}
memoir stuff -> memhfixc.sty

subfigure: \sub@label (add \@currentHref to the \newlabel

XR@addURL ??


\xspaceaddexceptions{\Hy@xspace@end,\hyper@linkend,\hyper@anchorend}
needed for new commands too?

\let\texttilde\~% for some languages

\url (to add links)


shipout hook: why pdftex different? (finish doc sets info)

\ltx@iffileloaded{hpdftex.def}{%
  \AtBeginShipout{%
    \Hy@FirstPageHook
    \global\let\Hy@FirstPageHook\ltx@empty
  }%
}{%
  \AtBeginShipoutFirst{%
    \Hy@FirstPageHook
  }%
}
\g@addto@macro\Hy@FirstPageHook{%
  \PDF@FinishDoc
  \global\let\PDF@FinishDoc\ltx@empty
}


the following can be stopped with implicit=false

\item (to add number for destination names)

\refstepcounter (to add anchor)

\appendix (to change the H-counter)

(fancyvrb) \FV@StepLineNo to add anchor

(lastpage/revtex4) \lastpage@putlabel (anchor)

(titlesec): \ttl@Hy@steplink, \ttl@Hy@PatchSaveWrite, \ttl@Hy@refstepcounter

(titletoc) \ttl@gobblecontents

(varioref) \vref@pagenum, \refstepcounter

(longtable) \LT@array? \LT@start (anchor)?

(amsmath): \equation, \endequation (anchors), \eqnarray, \endeqnarray, eqnarray*, endeqnarray*, subeqnarray, \endsubeqnarray, \subequations

(cleveref) \@thm

(amsthm) \@thm

==========
with hyperfootnotes:

\@footnotetext, \@footnotemark, \@mpfootnotetext 

(tabularx): \TX@endtabularx to add code for footnotes.

(longtable): \LT@p@ftntext for something with footnotes

(fancyvrb): \V@@footnotetext

\footref

\maketitle (something with foonotes)
========

\caption, \@caption: anchors

(float): \float@makebox for anchor

\bibcite, \@lbibitem, \@bibitem for anchors (unless natbib is loaded)

(revtex) \@CITEX, \@citex

(harvard) \HAR@citetoaux, \harvardcite, \harvarditem, \@harvarditem, \HAR@checkcitations

(chicago) \shortciteN, \citeN, \@citedata@opt, ...

(slides) \slide, \note, \overlay, ... for page anchors

\addcontentsline (destination name, bookmarks code, writes a \contentsline with four arguments)

\contentsline: uses fourth argument to build \Hy@tocdestname and to add a link in l@#1: 
      \csname l@#1\endcsname{%
        \hyper@linkstart{link}{\Hy@tocdestname}{#2}\hyper@linkend
      }{#3}%
      
\@definecounter, \@newctr, \@addtoreset: adds a setup for \theH<counter>      

(amsmath): \df@tag (for anchor)

\@seteqlabel (only with pdftex??)

(amsmath) \numberwithin for \theHcounter

\Gin@setfile (if hyperfigures, makes them a destination?)

\@wrindex, \@@wrindex: index stuff

A number of providing H-counter for various classes/styles: foils, seminar

(slidesec): a number of patches for hyperref stuff.

\@setref, \@pagesetref: to get link