 p: \pdfoutline\BKM@attr\BKM@action
                count\ifBKM@DO@open\else-\fi\BKM@x@childs
                {\BKM@DO@title}%
                
   \BKM@attr: attr{/F \BKM@DO@flags /C[\BKM@DO@color]}              

d:  \BKM@write{[}%
    \BKM@write{/Title(\BKM@DO@title)}%
    \ifnum\BKM@x@childs>\z@
      \BKM@write{/Count \ifBKM@DO@open\else-\fi\BKM@x@childs}%
    \fi
    \ifx\BKM@attr\@empty
    \else
      \BKM@write{\BKM@attr}%
    \fi
    \BKM@write{\BKM@action}%
    \BKM@write{/OUT pdfmark}%
    
  \BKM@attr: /F \BKM@DO@flags /C[\BKM@DO@color]
    
e.g. 
[ /Title(\376\377\000f\000i\000r\000s\000t\000p\000a\000g\000e)
/C[1 0 0]
/Action<</Subtype/Named/N/FirstPage>>
/OUT pdfmark 

[
/Title(\376\377\000a\000b\000c)
/Count -3
/Action/GoTo/Dest(section.1)cvn
/OUT pdfmark

   
 x: \special{pdf:%
        out
        [] OR [-] %depending is open or not
        \BKM@abslevel %current level. Must be equal or if larger one off to previous level
        <<
         /Title(\BKM@title)%

         /C[\BKM@color] %color, optional

         /F \BKM@FLAGS  % flags, optional

         \BKM@action    % action.
        >>%
      }%

[
/Title(\376\377\000a\000b\000c)
/Action/GoTo/Dest(subsection.1.1)cvn
/OUT pdfmark


p: goto page1{/Fit}
d: /Action/GoTo/Page 1 /View[/Fit]%
x: /Dest[@page1/Fit]

p: goto page\BKM@DO@page{/\BKM@DO@view} 
d: /Action/GoTo/Page \BKM@DO@page/View[/\BKM@DO@view]
x: /Dest[@page\BKM@page/\BKM@view]

uri:
p: user{<</S/URI/URI(\BKM@DO@uri)>>}
d: /Action<</Subtype/URI/URI(\BKM@DO@uri)>>%
x:  /A<</S/URI/URI(\BKM@uri)>>%

rawaction (should not be used, as clearly driver dependent
p:  user{<<\BKM@DO@rawaction>>}%
d:  /Action<<\BKM@DO@rawaction>>
x: /A<<\BKM@rawaction>>

Named
p: user{<</S/Named/N/\BKM@DO@named>>}
d: /Action<</Subtype/Named/N/\BKM@DO@named>>%
x: /A<</S/Named/N/\BKM@named>>%

p: \BKM@DefGotoNameAction\BKM@action\BKM@DO@dest ?
d: /Action/GoTo /Dest(\BKM@DO@dest)cvn%
x: /A<< /S/GoTo/D(\BKM@dest) >>%

for gotor:
p: /D[\BKM@DO@page/\BKM@DO@view]
d: /Page \BKM@DO@page /View[/\BKM@DO@view]%
x: {/D[\BKM@page/\BKM@view]}

p: /D(\BKM@DO@dest)
d: /Dest(\BKM@DO@dest)cvn%
x: /D(\BKM@dest)

gotor
p: user{ <</S/GoToR/F(\BKM@DO@gotor)\BKM@action>>%
d: /Action/GoToR/File(\BKM@DO@gotor)\BKM@action
x: /A<</S/GoToR/F(\BKM@gotor)\BKM@action>>%
 
dvipdfmx /Count is created automatically, so mainly the actions and options must be handled.
\bookmarks writes out the needed \special directly.

dvips writes out at the end a header file, jobname.out.ps, which is referenced with \special{header=#1}%
  count must done by the document, so every \bookmark creates a BKM@<id> command which stores {parent,level,count}.
  Childs update the count of their parent. Beside this every \bookmark writes the info with the destination (and more) to the aux-file as
  a command \BKM@entry. When the auxfile is read at end document these \BKM@entry writes the header file.

pdflatex,luatex: similar to dvips, only that \BKM@entry uses \pdfoutline directly.
 
    
    
Catalog
   /Outlines 56 0 R
             56 0 obj
             <<
            /Type /Outlines
            /First 24 0 R
            /Last 26 0 R
            /Count 3
            >>
            endobj
               24 0 obj
               <<
               /Title 25 0 R
               /A 23 0 R      --> << /S /GoTo /D (section.1) >>
               /Parent 56 0 R
               /Next 33 0 R
               /First 27 0 R
               /Last 30 0 R
               /Count -2
               >>
               endobj
                  27 0 obj
                 <<
                 /Title 28 0 R
                 /A 26 0 R
                 /Parent 24 0 R
                 /Next 30 0 R
                 >>
                 endobj

               33 0 obj
               <<
               /Title 34 0 R
               /A 32 0 R
               /Parent 56 0 R
               /Prev 24 0 R
               /Next 36 0 R
               >>
               endobj


Action types

/Type /Action /S XXXX /Next XXX

GoTo  : /D  (goto action is the same as /Dest)

GoToR : /F, /D, /NewWindow (goto remote)

GoToE : /F, /D, /NewWindow, /T (lots of subvalues) (goto embedded), pdf 1.6

GoToDp: (2.0) /Dp (goto part)

Launch:  /F, /NewWindows, (/Win etc deprecated in pdf 2.0)

Thread: /F, /D, /B

URI: /URI, /IsMap


Sound: /Sound, /Volume,/Synchronous,/Repeat,/Mix

Movie: deprecated in 2.0

Hide: /T, /H
 Set an annotation’s Hidden flag.

Named /N (e.g. /N /NextPage)

SubmitForm
(PDF 1.2) Send data to a uniform resource locator.
“Submit-Form Actions” on page 703
ResetForm
(PDF 1.2) Set fields to their default values.
“Reset-Form Actions” on page 707
ImportData
(PDF 1.2) Import field values from a file.
“Import-Data Actions” on page 708
JavaScript
(PDF 1.3) Execute a JavaScript script.
“JavaScript Actions” on page 709
SetOCGState /State , /PreserveRB

Rendition

Trans
(PDF 1.5) Updates the display of a document, using a transition dictionary.
“Transition Actions” on page 670
GoTo3DView
